MultiATSM package - General Guidelines

### 2023-08-16 inst/doc/my-vignette.Rmd
version 0.3.3

In what follows, the focus will be devoted to describe the various pieces that are necessary to implement ATSMs. Specifically, Section \@ref(S:SectionData) describes the data-set available at this package and the set of functions that are useful to retrieve data from Excel files. Section \@ref(S:SectionInputs) exposes the necessary inputs that have to be specified by the user. In Section \@ref(S:SectionEstimation) the estimation procedure is detailed. Section \@ref(S:SectionReplication) shows how to use the *MultiATSM* package to estimate ATSMs from scratch as well as to closely reproduce some empirical features of academic papers. To ease the exposition, throughout the next sections, the database used by CM (2021) is employed to illustrate the use of the various functions available in this package.

@@

The GVAR factors database casts country-specific lists of all the factors that are used in the estimation of each country's VARX(1,1,1). A specific function for computing the star risk factors is detailed in the Section \@ref(S:SectionGVAR). 
```{r}
data('CM_Factors_GVAR')
```

@@

- **Model type**: a string-vector containing the label of the model to be estimated (as per described in Table \@ref(tab:ModFea)); 
- **Frequency of the data**: a string-vector that specifies the frequency of the time series data. The available options are: *"Annually"*, *"Quarterly"*, *"Monthly"*, *"Weekly"*, *"Daily Business Days"*, and *"Daily All Days"*;
- **Economies**: a string-vector containing the names of the economies which are part of the economic system;
- **Number of spanned factors (N)**: number of country-specific spanned factors. In this package, *N* is assumed to be equal across countries;
- **Global variables**: a string-vector containing the labels of the global unspanned factors used in the model estimation;
- **Domestic variables**: a string-vector containing the labels of the domestic unspanned factors used in the model estimation;
- **Bond yield set of maturities**: a numerical-vector containing the maturities of the yields used in the model estimation. The units of this vector must be expressed in *years*. Alternatively, the user can rely on the *Maturities* routine to extract the bond maturities from the time series of bond yields (see the documentation of the *Maturities* function). In this case, the user must specify the **unit in which the label of the bond yields** are expressed, e.g. if the label of the 10-years bond is selected as *Y120M* (*Y10y*), then the option *Month* (*Year*) must be selected;
- **Factor labels**: a string-list based which contains the labels of all the variables present in the model. This input is particularly useful for constructing the plots. The routine *LabFac* provides the desired list of factor labels;
- **Stationarity constraint under the $Q$-dynamics**: a binary variable which takes value equal to "1" if the user wishes the largest eigenvalue under the $Q$-measure to be strictly smaller than 1. Otherwise, this same variable must be set to "0";
- **Output label**: a single element string-vector which contains the of the name that appears in the file name that stores the model outputs.

One possible example of the basic model inputs is
```{r}
ModelType <- "JPS"
DataFrequency <- "Monthly"
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
N <- 3 # number of spanned factors per country
GlobalVar <- c("GBC", "CPI_OECD") # Global variables
DomVar <- c("Eco_Act", "Inflation") # Domestic variables 
mat <- c(0.25, 0.5, 1, 3, 5, 10) # vector of maturities
FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType) 
StationarityUnderQ <- 0 # 1 = set stationarity condition under  the Q; 0 = no stationarity condition
OutputLabel <- "Model_demo" # output label
```

### GVARinputs and JLLinputs {#S:SectionGVARJLLinputs}
Some additional inputs are required if the user intends to estimate the GVAR-ATSM or JLL-ATSM related setups. For a convenient use, the extra outputs should be stored in a separate list for each model. The GVAR model list of inputs must contain:

@@

@@

- Model fit of the bond yields;
- Impulse response functions (IRFs);
- Forecast error variance decompositions (FEVDs);
- Generalized impulse response functions (GIRFs);
- Generalized forecast error variance decompositions (GFEVDs).

@@

For the graphical outputs, the user must indicate the desired types of graphs in a string-based vector:
```{r}
DesiredGraphs <- c("Fit", "GIRF", "GFEVD") # Available options are: "Fit", "IRF", "FEVD", "GIRF", "GFEVD"
```

@@

- **methodBS**: the desired bootstrap procedure. Available options are *(i)* standard residual bootstrap ('bs'); *(ii)* wild bootstrap ('wild') and block bootstrap ('block'). If the latest is selected, then the block length must be indicated;
- **ndraws**:  the number of bootstrap draws;
- **pctg**: the confidence level expressed in percentage points.

@@

# Model Estimation {#S:SectionEstimation}
Having gathered bond yields and other economic time series data, the ATSM estimation is carried out in three steps. Firstly, one needs to provide the number of country-specific spanned factors to be included in the global ATSM. Secondly, it is decided on the form of the risk factor dynamics under the $P$-measure. Finally, the user needs to select general features for the model optimization.

@@

```{r}
data("CM_Trade")
t_First <- "2006"
t_Last <-  "2019"
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
type <- "Sample Mean"
W_gvar <- Transition_Matrix(t_First, t_Last, Economies, type, DataPath = NULL, TradeFlows)
```

@@

A separate routine is provided for computing the foreign-specific factors (also commonly referred as the star variables) used in the estimation of the VARX models.
```{r}
data('CM_Factors')
StaFac <- StarFactors(RiskFactors, Economies, W_gvar)
```

@@

# Examples of full implementation of ATSMs {#S:SectionReplication}
This section presents some examples on how to use the *MultiATSM* package to fully implement ATSMs. Further on, the functions of this package are used to replicate some of the results presented in the original papers of JPS (2014) and CM (2021).

## General template
```{r, eval=FALSE}
###############################################################################################################
#################################### USER INPUTS ##############################################################
###############################################################################################################
# A) Load database data
data("CM_Factors")
data('CM_Factors_GVAR')
data('CM_Trade')
data('CM_Yields')

# B) Decide on general model inputs
ModelType <- "JLL original" # available options are "JPS", "JPS jointP", "GVAR sepQ", "VAR jointQ", "GVAR jointQ", "JLL original", "JLL NoDomUnit", "JLL jointSigma".

StationarityUnderQ <- 1 # Wish to impose stationary condition for the eigenvalues of each country: YES: 1, NO:0

@@

OutputLabel <- "Test" # label of the model for saving the file
DataFrequency <- "Monthly" # Frequency of the data
UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"

# B.1) Decide on specific model inputs
#################################### GVAR-based models #########################################################
if (ModelType == 'GVAR sepQ' || ModelType == 'GVAR jointQ'){
  # Transition matrix inputs:
  t_First <- "2006" # First year of the sample
  t_Last <-  "2019" # Last year of the sample
  W_type <- 'Sample Mean' # Method to compute the transition matrix
  # VARX input:
  VARXtype <- "unconstrained"
}

#################################### JLL-based models #########################################################
if (ModelType =="JLL original" || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma"){
  DominantCountry <- "China" # (i) one of the countries of the system or (ii) "None"
}
################################################################################################################

# C) Decide on Settings for numerical outputs
Horiz <- 50
DesiredGraphs <- c("IRF", "FEVD") #c("Fit", "IRF", "FEVD", "GIRF", "GFEVD")
WishGraphRiskFac <- 0
WishGraphYields <- 1
WishOrthoJLLgraphs <- 0

# D) Bootstrao settings
WishBootstrap <- 1 #  YES: 1; No = 0.
Bootlist <- list()
Bootlist$methodBS <- 'bs' # (i) 'bs' ; (ii) 'wild'; (iii) 'block'
Bootlist$BlockLength <- 4 # necessary input if one chooses the block bootstrap
Bootlist$ndraws <-  30
Bootlist$pctg   <-  95 # confidence level

@@

###############################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE #############################################
###############################################################################################################

# 2) Minor preliminary work
C <- length(Economies)
FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType) # Generate the set of labels
mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields) # Vector of maturities expressed in years
# (alternatively, it can be completed manually)
ZZ <- RiskFactors
# 2.1) Build the GVARinputs
if (ModelType == 'GVAR sepQ' || ModelType == 'GVAR jointQ'){
  GVARinputs <- list()
  GVARinputs$Economies <- Economies
  GVARinputs$GVARFactors <- FactorsGVAR
  GVARinputs$VARXtype <- VARXtype
  GVARinputs$Wgvar <- Transition_Matrix(t_First, t_Last, Economies, W_type, DataPath = NULL, Data = TradeFlows)
} else { GVARinputs <- NULL }

# 2.2) Build the JLLinputs
if (ModelType =="JLL original" || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma"){
  JLLinputs <- list()
  JLLinputs$Economies <- Economies
  JLLinputs$DomUnit <- DominantCountry
  JLLinputs$WishSigmas <- 1 # Sigma matrix is estimated within the "InputsForMLEdensity" function
  JLLinputs$SigmaNonOrtho <- NULL
  JLLinputs$JLLModelType <- ModelType
}else{
  JLLinputs <- NULL
}

# 3) Prepare the inputs of the likelihood function
ModelParaList <- list()
for (i in 1:C){
  if (( ModelType == "GVAR jointQ" || ModelType == "VAR jointQ" || ModelType == "JLL original"
        || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma" )   & i >1 ){break} # the break avoids the models with joint estimation under the Q to be estimated C times

  # 3.1) Compute the inputs that go directly into the log-likelihood function
  ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency, JLLinputs,   GVARinputs)

  # 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
  K1XQ <- ATSMInputs$K1XQ
  if (ModelType == "JLL original" || ModelType == "JLL NoDomUnit" ){ SSZ <- NULL} else{SSZ <- ATSMInputs$SSZ}

@@

  # 6) Optimization of the model
  if (ModelType == 'JPS' || ModelType == 'JPS jointP' || ModelType == "GVAR sepQ"){
    ModelParaList[[ModelType]][[Economies[i]]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs, GVARinputs)$Summary
  }else{ ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs, GVARinputs)$Summary}

}

# 7) Numerical and graphical outputs
InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ,                  WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, WishBootstrap, Bootlist, WishForecast, ForecastList)
# A) Fit, IRF, FEVD, GIRF and GFEVD
NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)

# B) Bootstrap
Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs, FactorLabels, DataFrequency, varargin, JLLinputs, GVARinputs)

# C) Out-of-sample forecasting
Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies, DataFrequency, JLLinputs, GVARinputs)

@@

## Paper replications
### Joslin, Priebisch and Singleton (2014)
To the best of my knowledge, the original data-set used in @JoslinPriebschSingleton2014 is not publicly available. Nevertheless, there exists an alternative source of data that closely matches JPS (2014) in terms of the construction method and the sample period. This database is constructed by @BauerRudebusch2017 (BR, 2017) and is available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/). In that paper, BR (2017) investigate whether macro-finance term structure models better suit the unspanned macro risk framework by JPS (2014) or the earlier traditional spanned settings as the one by @AngPiazzesi2003JME. Accordingly, BR (2017) intend to replicate some of the empirical results reported in JPS (2014). The R-code used by BR (2017) is also available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/).

Supported by BR (2017)'s data-set, the code below uses the *MultiATSM* package to estimate the key model parameters from the ATSM along the lines of JPS (2014).

```{r, eval=FALSE} 
###############################################################################################################
#################################### USER INPUTS ##############################################################
###############################################################################################################
# A) Load database data
data("BR_jps_gro_R3")

# B) Decide on general model inputs
ModelType <- "JPS"

StationarityUnderQ <- 0 # Wish to impose stationary condition for the eigenvalues of each country: YES: 1, NO:0

Economies <- c("U.S.") # Names of the economies from the economic system
GlobalVar <- c()
DomVar <- c("GRO", "INF") # Country-specific variables

DataFrequency <- "Monthly" # Frequency of the data
UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"

mat <- c(0.25, 0.5, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10) # maturities expressed in years

###############################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE #############################################
###############################################################################################################
# 2) Minor preliminary work
C <- length(Economies)
M <- length(DomVar)
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType) # Generate the set of labels
Yields <- t(BR_jps_out$Y)
W <- BR_jps_out$W[1:N,] # Use the Wpca matrix from BR (2017)
SpaFac <- W%*%Yields
rownames(SpaFac) <- FactorLabels$Tables$U.S.[-(1:M)]
ZZ <- rbind(t(BR_jps_out$M.o), SpaFac) # Complete set of risk factors

# 3) Prepare the inputs of the likelihood function
for (i in 1:C){
# 3.1) Compute the inputs that go directly into the log-likelihood function
ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency)
# 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
K1XQ <- ATSMInputs$K1XQ
SSZ <- ATSMInputs$SSZ
# 3.3) Adjust the inputs which are funtion of the W matrix
ATSMInputs$Wpca <- W
ATSMInputs$We <- t(pracma::null(W))
ATSMInputs$WpcaFull <- rbind(ATSMInputs$Wpca, ATSMInputs$We)
ATSMInputs$PP <- SpaFac

# 6) Optimization of the model
ModelPara <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType)$Summary
}
```

The tables below compare the ATSM parameter estimates generated from BR (2017) and the *MultiATSM*. Overall, one can note that the differences in the estimates are economically modest.

```{r, echo= FALSE}
options(scipen = 100) # eliminate the scientific notation
data("BR_jps_gro_R3")
data("JPSrep")

RowsQ <- c("$r0$", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_3$" )
TableQ <- data.frame(matrix(NA, ncol = 0, nrow =length(RowsQ)))
row.names(TableQ) <- RowsQ

PackageQ<- c(ModelPara$ests$r0, diag(ModelPara$ests$K1XQ))
BRq <- c(BR_jps_out$est.llk$rho0.cP, diag(BR_jps_out$est.llk$KQ.XX))
TableQ$MultiATSM <- PackageQ
TableQ$'BR (2017)' <- BRq

TableQ <- round(TableQ, digits = 5)

kableExtra::kbl(TableQ, align = "c", caption = "$Q$-dynamics parameters") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 14) %>%
  kableExtra::footnote(general = " $\\lambda$'s are the eigenvalues from the risk-neutral feedback matrix and $r0$ is the long-run mean of the short rate under Q.")
```

```{r, echo= FALSE}
data("BR_jps_gro_R3")
data("JPSrep")

RowsP <- c("PC1", "PC2", "PC3", "GRO", "INF")
ColP <- c(" ", RowsP)

# 1) K0Z and K1Z
# Bauer and Rudebusch coefficients
TablePbr <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
row.names(TablePbr) <- RowsP
colnames(TablePbr) <- ColP

TablePbr[[ColP[1]]] <- BR_jps_out$est.llk$KP.0Z
for(j in 1:length(RowsP) ){TablePbr[[RowsP[j]]] <- BR_jps_out$est.llk$KP.ZZ[,j]}

TablePbr <- round(TablePbr, digits = 5)

# MultiATSM coefficients
TablePMultiATSM <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
row.names(TablePMultiATSM) <- RowsP
colnames(TablePMultiATSM) <- ColP

IdxVar <- c(3:5, 1:2) # indexes to flip the order of the spanned and unspanned factors
TablePMultiATSM[[ColP[1]]] <- ModelPara$ests$K0Z[IdxVar]
ModelPara$ests$K1Z <- ModelPara$ests$K1Z[IdxVar,IdxVar]
for(j in 1:length(RowsP) ){ TablePMultiATSM[[RowsP[j]]] <- ModelPara$ests$K1Z[,j]}

TablePMultiATSM <- round(TablePMultiATSM, digits = 5)
TableP <- rbind(TablePbr,TablePMultiATSM)
row.names(TableP) <- c(RowsP,paste(RowsP," ",sep="")) # trick to avoid rows to have the same name

kableExtra::kbl(TableP, align = "c", caption = "$P$-dynamics parameters") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 14) %>%
  kableExtra::add_header_above(c(" "= 1, "K0Z"=1, "K1Z" = 5), bold = T) %>%
  kableExtra::pack_rows("BR (2017)", 1, 5) %>%
  kableExtra::pack_rows("MultiATSM", 6, 10) %>%
  kableExtra::footnote(general = " $K0Z$ is the intercept and $K1Z$ is feedback matrix from the $P$-dynamics.")
```

```{r, echo= FALSE}
data("BR_jps_gro_R3")
data("JPSrep")

se <- data.frame(BR_jps_out$est.llk$sigma.e, ModelPara$ests$se )
rownames(se) <- "se"
colnames(se) <- c("MultiATSM","BR (2017)")
se <- round(se, digits = 7)

kableExtra::kbl(se, align = "c", caption ="Portfolio of yields with errors") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 14) %>%
  kableExtra::footnote(general = " $se$ is the standard deviation of the portfolio of yields observed with errors.")
```

It is relevant to highlight that the script above makes use of the principal component weights provided by BR (2017). Such a matrix is simply a scaled-up version of the one provided by the function *pca_weights_one_country* of this package. Accordingly, despite the numerical differences on the weight matrices, both methods generate time series of spanned factors which are perfectly correlated. Another difference between the two approaches relates to the construction form of the log-likelihood function (llk): while in the BR (2017) code the llk is expressed in terms of a portfolio of yields, the *MultiATSM* package generates this same input directly as a function of observed yields (i.e. both procedures lead to equivalent llk up to the Jacobian term).

### Candelon and Moura (2021)
```{r, eval=FALSE}
###############################################################################################################
#################################### USER INPUTS ##############################################################
###############################################################################################################
# A) Load database data
data("CM_Factors")
data('CM_Factors_GVAR')
data('CM_Trade')
data('CM_Yields')

# B) Decide on general model inputs
ModelType <- "GVAR jointQ" # Set "GVAR jointQ" for the GVAR-ATSM and "JLL original" for the JLL-ATSM

StationarityUnderQ <- 0

Economies <- c("China", "Brazil", "Mexico", "Uruguay") 
GlobalVar <- c("GBC", "CPI_OECD") # Global Variables
DomVar <- c("Eco_Act", "Inflation") # Country-specific variables

OutputLabel <- "CM_2021" # label of the model for saving the file
DataFrequency <- "Monthly" # Frequency of the data
UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"

>>>>

In what follows, the focus will be devoted to describe the various pieces that are necessary to implement ATSMs. Specifically, Section \@ref(S:SectionData) describes the data-set available at this package and the set of functions that are useful to retrieve data from Excel files. Section \@ref(S:SectionInputs) exposes the necessary inputs that have to be specified by the user. In Section \@ref(S:SectionEstimation) the estimation procedure is detailed. Section \@ref(S:SectionReplication) shows how to use the *MultiATSM* package to estimate ATSMs from scratch. To ease the exposition, throughout the next sections, the database used by CM (2021) is employed to illustrate the use of the various functions available in this package.

@@

The GVAR factors database casts country-specific lists of all the factors that are used in the estimation of each country's VARX(1,1,1). The function titled *DatabasePrep* provides assistance to structure the data in a similar fashion to this form. Moreover, a specific function for computing the star risk factors is detailed in the Section \@ref(S:SectionGVAR). 
```{r}
data('CM_Factors_GVAR')
```

@@

- **Model type**: a string-vector containing the label of the model to be estimated (as per described in Table \@ref(tab:ModFea)); 
- **Model parameters corrected for bias**: following the bias correction method described in @BauerRudebuschWu2012 (BRW, 2012), this variable assumes value equal to "1" if the user wishes to implement the framework of BRW (2012) and "0", otherwise;
- **Frequency of the data**: a string-vector that specifies the frequency of the time series data. The available options are: *"Annually"*, *"Quarterly"*, *"Monthly"*, *"Weekly"*, *"Daily Business Days"*, and *"Daily All Days"*;
- **Economies**: a string-vector containing the names of the economies which are part of the economic system;
- **Number of spanned factors (N)**: number of country-specific spanned factors. In this package, *N* is assumed to be equal across countries;
- **Global variables**: a string-vector containing the labels of the global unspanned factors used in the model estimation;
- **Domestic variables**: a string-vector containing the labels of the domestic unspanned factors used in the model estimation;
- **Bond yield set of maturities**: a numerical-vector containing the maturities of the yields used in the model estimation. The units of this vector must be expressed in *years*. Alternatively, the user can rely on the *Maturities* routine to extract the bond maturities from the time series of bond yields (see the documentation of the *Maturities* function). In this case, the user must specify the **unit of time in which the label of the bond yields** are expressed, e.g. if the label of the 10-years bond is selected as *Y120M* (*Y10y*), then the option *Month* (*Year*) must be selected;
- **Stationarity constraint under the $Q$-dynamics**: a binary variable which takes value equal to "1" if the user wishes the largest eigenvalue under the $Q$-measure to be strictly smaller than 1. Otherwise, this same variable must be set to "0";
- **Computation of forward premia**: this is another binary variable which takes the value equal to "1" if the user wishes the computation of the forward premia or "0", otherwise. If the this variable is set to "1", then the user must specify a two-element numerical vector containing the maturties associated with the starting and the ending date of the loan;
- **Output label**: a single element string-vector which contains the of the name that appears in the file name that stores the model outputs.

One possible example of the basic model inputs is
```{r}
ModelType <- "JPS"
BiasCorrection <- 1 # 1 = estimate model with bias correction; 0 = estimate model without bias correction
DataFrequency <- "Monthly"
UnitMatYields <- "Month" # Available options are: "Month" or "Year".
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
N <- 3 # number of spanned factors per country
GlobalVar <- c("GBC", "CPI_OECD") # Global variables
DomVar <- c("Eco_Act", "Inflation") # Domestic variables 
mat <- c(0.25, 0.5, 1, 3, 5, 10) # vector of maturities
StationarityUnderQ <- 0 # 1 = set stationarity condition under  the Q; 0 = no stationarity condition
WishForwardPremia <- 1 # Wish to estimate the forward premia: 1= Yes; 0= No
FPmatLim <- c(60,120) # maturty of the loan starts in 60 months and ends in 120 months in the future
OutputLabel <- "Model_demo" # output label
```

### GVARinputs and JLLinputs {#S:SectionGVARJLLinputs}
Some additional inputs are required if the user intends to estimate the GVAR-ATSM or JLL-ATSM related setups. The extra inputs should be stored in separate lists for each model for convenience. To assist the user in executing this task, one can use the *ListModelInputs* function. Otherwise, the *GVARinputs* and *JLLinputs* constructs can be formed manually as specified below.

To GVAR models, the set of supplementary inputs include:

@@

### BRWinputs {#S:SectionBRWinputs}
In an influential paper, BRW (2012) have shown that the estimates from traditional ATSMs can be severely biased due to the typical small-sample size used in these studies. As result, such models may produce 
unreasonably stable expected future short-term interest, therefore distorting the term premia estimates for long-maturity bonds. To circumvent this problem, BRW (2012) propose a bootstrap-based method that relies on indirect inference estimation that corrects for the bias present in previous works.

From its version $0.3.1$, the *MultiATSM* package accomodates the framework proposed by BRW (2012) to each one of its ATSM types. Some additional inputs must be specified, if the user intends to perform the model estimation along the lines of BRW (2012). They are:

- **Mean or median of OLS estimates**: the user must decide whether to compute the mean or median of the OLS estimates after each bootstrap iteration. To compute the mean (median), the user must set *TRUE* (*FALSE*);
- **Number of iteration**: total amount of iterations used in the stochatic approximation algorithm after burn-in;
- **Number of burn-in iteration**:  quantity of the iterations to be discarded in the first stage of the bias-correction estimation process. The recommended number is $15\%$ of the total number of iterations;
- **Number of bootstrap samples**:  quantity of simulated samples used in each burn-in or actual iteration
- **Adjustment parameter**: this parameter controls the degree of shrinkage between the difference in the OLS estimates and the BRW (2012)'s bootstrap-based one after each iteration. The value of this parameter is fixed and must lie in the interval $]0;1[$;
- **Perform closeness check**: flag whether the user wishes to check the root mean square distance between the model estimates produced by the bias-correction method and those generated by OLS. Default is set to *TRUE*;
- **Number of bootstrap samples used in the closeness check**: default is set to 100,000;

```{r}
flag_mean <- TRUE # TRUE = compute the mean; FALSE = compute the median
gamma <- 0.2 # Adjustment parameter
N_iter <- 500 # Number of iteration to be conserved 
N_burn <- N_iter*0.15  # Number of iteration to be discarded 
B <- 50 # Number of bootstrap samples 
checkBRW <- 1 # Closeness check
B_check <- 1000 # Number of bootstrap samples used in the closeness check
```

@@

- Model fit of the bond yields;
- Impulse response functions (IRFs);
- Forecast error variance decompositions (FEVDs);
- Generalized impulse response functions (GIRFs);
- Generalized forecast error variance decompositions (GFEVDs);
- Bond yields decomposition into their expected component and term premia.

@@

For the graphical outputs, the user must indicate the desired types of graphs in a string-based vector:
```{r}
DesiredGraphs <- c("Fit", "GIRF", "GFEVD") # Available options are: "Fit", "IRF", "FEVD", "GIRF", 
                                          # "GFEVD", "TermPremia".
```

@@

- **methodBS**: the desired bootstrap procedure. Available options are *(i)* standard residual bootstrap ('bs'); *(ii)* wild bootstrap ('wild'), and block bootstrap ('block'). If the latest is selected, then the block length must be indicated;
- **ndraws**:  the number of bootstrap draws;
- **pctg**: the confidence level expressed in percentage points.

@@

# Model Estimation {#S:SectionEstimation}
Having gathered bond yields and other economic time series data, the ATSM estimation is carried out in three steps. Firstly, one needs to provide the number of country-specific spanned factors to be included in the global ATSM. Secondly, it is decided on the form of the risk factor dynamics under the $P$-measure. Finally, the user needs to select the general features for the model optimization.

@@

```{r}
data("CM_Trade")
t_First <- "2006"
t_Last <-  "2019"
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
type <- "Sample Mean"
W_gvar <- Transition_Matrix(t_First, t_Last, Economies, type, DataPath = NULL, TradeFlows)
print(W_gvar)
```

Starting from package version $0.3.3$, the GVAR setups can be estimated incorporating time-varying interdependence weights. This entails that the star factors are constructed by applying trade flow weights from the specific year to adjust the risk factors of the corresponding year. To employ this feature, users need to specify the transition matrix type as *Full Sample* and select the **same specific date (year)** for both the initial and final years within the transition matrix. This implies that the trade weights of this year are used for solving (link matrices) the GVAR system. Taking reference from the variable names in the previous example, an illustration could be set as follows:

```{r}
t_First <- "2019"
t_Last <-  "2019"
type <- "Full Sample"
```

@@

A separate routine is provided for computing the foreign-specific factors (also commonly referred to as the star variables) used in the estimation of the VARX models.
```{r}
data('CM_Factors')
StaFac <- StarFactors(RiskFactors, Economies, W_gvar)
```

@@

# Examples of full implementation of ATSMs {#S:SectionReplication}
This section presents some examples on how to use the *MultiATSM* package to fully implement ATSMs. Further on, the functions of this package are used to replicate some of the results presented in the original papers of JPS (2014) and CM (2021). See the [Paper Replication Vignette](Paper-Replications.html) of this package for a detailed reproduction.

## General template
```{r, eval=FALSE}
########################################################################################################
#################################### USER INPUTS #######################################################
########################################################################################################
# A) Load database data
data("CM_Factors")
data('CM_Factors_GVAR')
data('CM_Trade')
data('CM_Yields')

# B) Decide on general model inputs
ModelType <- "VAR jointQ" # available options are "JPS", "JPS jointP", "GVAR sepQ", "VAR jointQ", 
                          #"GVAR jointQ", "JLL original", "JLL NoDomUnit", "JLL jointSigma".

StationarityUnderQ <- 0 # Wish to impose stationary condition for the eigenvalues of each country: 
                         #YES: 1,NO:0
BiasCorrection <- 1 # Wish to estimate the model with the bias correction method of BRW (2012): 
                    #YES: 1, NO:0

WishForwardPremia <- 1 # Wish to estimate the forward premia: YES: 1, NO:0
FPmatLim <- c(60,120) #  If the forward premia is desired, then choose the Min and max maturities of the    
                        # forward premia. Otherwise set NA

@@

OutputLabel <- "Test" # label of the model for saving the file
DataFrequency <- "Monthly" # Frequency of the data
UnitMatYields <- "Month" # bond yields time-unit. Available options are "Month" or "Year"

# B.1) Decide on specific model inputs
#################################### GVAR-based models ##################################################
t_First_Wgvar <- "2004" # First year of the sample (transition matrix)
t_Last_Wgvar <-  "2019" # Last year of the sample (transition matrix)
W_type <- 'Sample Mean' # Method to compute the transition matrix
VARXtype <- "unconstrained" # (i) "unconstrained" or (ii) ""constrained" (VARX)
#################################### JLL-based models ###################################################
DomUnit <- "China"
WishSigmas <- 1 # Sigma matrix is estimated within the "InputsForMLEdensity" function
SigmaNonOrtho <- NULL
JLLModelType <- ModelType
###################################### BRW inputs  ######################################################
flag_mean <- TRUE # TRUE = compute the mean; FALSE = compute the median
gamma <- 0.2 # Adjustment parameter
N_iter <- 500 # Number of iteration to be conserved 
N_burn <- N_iter*0.15  # Number of iteration to be discarded 
B <- 50 # Number of bootstrap samples 
checkBRW <- 1
B_check <- 1000 # 
#########################################################################################################

# C) Decide on Settings for numerical outputs
Horiz <- 50
DesiredGraphs <- c("Fit", "IRF", "FEVD") # "Fit", "IRF", "FEVD", "GIRF", "GFEVD", "TermPremia"
WishGraphRiskFac <- 0
WishGraphYields <- 1
WishOrthoJLLgraphs <- 0

# D) Bootstrao settings
WishBootstrap <- 1 #  YES: 1; No = 0.
Bootlist <- list()
Bootlist$methodBS <- 'bs' # (i) 'bs' ; (ii) 'wild'; (iii) 'block'
Bootlist$BlockLength <- 4 # necessary input if one chooses the block bootstrap
Bootlist$ndraws <-  5
Bootlist$pctg   <-  95 # confidence level

@@

#########################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE #######################################
#########################################################################################################

# 2) Minor preliminary work
C <- length(Economies)
FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType) # Generate the set of labels
ZZ <- RiskFactors
if(any(ModelType == c("GVAR sepQ", "GVAR jointQ"))){
Data <- list()
Data$GVARFactors <- FactorsGVAR}
mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields)

# 2.1) Generate GVARinputs, JLLinputs and BRWinputs
ModInputs <- ListModelInputs(ModelType, Data, Economies, VARXtype, t_First_Wgvar, t_Last_Wgvar, W_type,
                             DomUnit, WishSigmas, SigmaNonOrtho, BiasCorrection, flag_mean, gamma, N_iter,
                             N_burn, B, checkBRW, B_check)

GVARinputs <- ModInputs$GVARinputs
JLLinputs <- ModInputs$JLLinputs
BRWinputs <- ModInputs$BRWinputs

# 3) Prepare the inputs of the likelihood function
ModelParaList <- list()
for (i in 1:C){
  if (( any(ModelType ==c("GVAR jointQ", "VAR jointQ","JLL original", "JLL NoDomUnit","JLL jointSigma"))) 
      & i >1 ){break}

  # 3.1) Compute the inputs that go directly into the log-likelihood function
  ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency,
                                    JLLinputs, GVARinputs, BRWinputs)

  # 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
  K1XQ <- ATSMInputs$K1XQ
  if (any(ModelType == c("JLL original", "JLL NoDomUnit"))){ SSZ <- NULL} else{SSZ <- ATSMInputs$SSZ}

@@

  # 6) Optimization of the model
  if (ModelType == 'JPS' || ModelType == 'JPS jointP' || ModelType == "GVAR sepQ"){
ModelParaList[[ModelType]][[Economies[i]]] <- Optimization(f, tol, varargin, FactorLabels,
                                                           Economies, ModelType)$Summary
  }else{
ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType,
                                           JLLinputs, GVARinputs)$Summary}
}

# 7) Numerical and graphical outputs
InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ,
                                     UnitMatYields, WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs,
                                     WishForwardPremia, FPmatLim, WishBootstrap, Bootlist, WishForecast,
                                     ForecastList)
# A) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)

# B) Bootstrap
Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs,
                       FactorLabels, DataFrequency, varargin, JLLinputs, GVARinputs, BRWinputs)

# C) Out-of-sample forecasting
Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
                            DataFrequency, JLLinputs, GVARinputs, BRWinputs)

@@
